---
title: "Example: Carbon tracking with Hector"
author: "Leeya Pressburger"
date: "2021-08-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example: Carbon tracking with Hector}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This vignette will walk through Hector's carbon tracking feature and show how to
turn on tracking, retrieve the tracking data, and display the results. Additionally, the vignette demonstrates an application of the carbon tracking feature and shows how to calculate the model's airborne fraction.

Hector's carbon tracking feature helps to trace the movement of carbon between 
various pools as the model runs. At the user-chosen start date, each pool is
deemed to be composed of 100% carbon from itself, i.e. at that moment in time 
the deep ocean pool is 100% composed of carbon from the deep pool. 

As the model moves forward in time from the start date, Hector tracks each pool's fractional composition -- i.e. the sources of its current carbon. 
When the run is finished, this data can be read into R and then analyzed.

## Setup

First, let's load the packages we will need.

``` {r hector.init}
library(hector)
library(ggplot2)
```

## Carbon tracking

Now, we will read in an .ini file and create a new core. In this example, we 
will use RCP 4.5.

In order to use the carbon tracking feature, we must use `setvar` and `fetchvars`. ^[Command line users: to set a tracking start date, change the `trackingDate=9999` line in the `[core]` section of an INI scenario file.]
First, let's look at the `trackingDate` prior to setting it to any value. 

``` {r setup.core}
inifile <- file.path(system.file('input', package = 'hector'),
                     'hector_rcp45.ini')
core <- newcore(inifile)

fetchvars(core, NA, TRACKING_DATE())
```

This returns the variable `trackingDate` with a `value` of 9999. This default 
value does not turn carbon tracking on (because the model never reaches year 9999).

Now, we will set a `trackingDate`, check that it worked, and run a core. 
Let's begin tracking in 1750. On that date, the model will mark all carbon in
each of its pools as originating in that pool: the atmosphere will be 100%
'atmosphere-origin' C, the soil 100% 'soil-origin' C, etc.

``` {r set.date}
setvar(core, NA, TRACKING_DATE(), 1750, "(unitless)")
fetchvars(core, NA, TRACKING_DATE())

invisible(reset(core))
run(core)
```

At this point, the model has been run. To retrieve the carbon tracking data, we
use the command `get_tracking_data()`. This returns a data frame which gives the 
current state of the pool, and fractions of where the carbon is sourced from, at
each model time point after `trackingDate`.

``` {r hector.tracking}
tdata <- get_tracking_data(core)
head(tdata)
```

The `tdata` dataset now contains information on year (note that the first year is the start-tracking date we asked for), the model component,
the name of the pool within that component, the amount of 
carbon in the pool, and the fraction (0-1) of each source for that pool. 
Importantly, the source fractions for a given pool and year will always sum to 1.

Side note: the names in the carbon tracking data (`pool_name`, `pool_value`,
and `pool_units`) differ from those used in the model's standard stream output
(which are `variable`, `value`, and `units` respectively). In contrast, `year` and `component` are named identically in the two output datasets.

## Results

By using `ggplot`, we can create time series graphs of how the origin of one
or more carbon pools in the model changes over time. 

``` {r hector.plot1, fig.width = 7, fig.height = 5}
ggplot(tdata, aes(year, source_fraction, fill = source_name)) +
  geom_area() +
  facet_wrap(~pool_name) +
  theme(axis.text = element_text(size = 7)) +
  labs(x = "Year",
       y = "Source Fraction",
       title = "Source Percentage Change Over Time")
```

Let's take a look at these graphs. Note that the `earth_c` pool is always 100% 
`earth_c`: there are no inflows to this pool in RCP 4.5 (in contrast, in
RCP 2.6 there is direct-air carbon capture and storage, so the `earth_c` pool
origins will change). We see that around 1800, human emissions (from `earth_c`) 
begin to appear in pools such as `atmos_c` or `soil_c_global`. 
Between 1900-2000, human emissions take off and become major sources by 2100. 

Let's examine just one pool: 

``` {r hector.atm, fig.width = 6, fig.height = 4}
atmos <- subset(tdata, pool_name == "atmos_c")

ggplot(atmos, aes(year, source_fraction, fill = source_name)) +
  geom_area() +
  facet_wrap(~pool_name) +
  labs(x = "Year",
       y = "Source Fraction",
       title = "Source Percentage Change Over Time - Atmosphere")
```

This allows us to better examine the contributions by source to the atmosphere. 

## Example: Airborne fraction calculation

We can also use the tracking feature to calculate the airborne fraction for a given point in time, or how much carbon emitted by humans remains in the atmosphere as a fraction of total emissions as opposed to being absorbed by the land or ocean. 

First, we can define two functions. The first will run a core and turn on tracking to return a data frame with results, including a column for the source amount by pool (as opposed to the source fraction). The second function will perform the airborne fraction calculation by finding the amount of `earth_c` (human emissions) in the `atmos_c` (atmosphere) pool at the start and end date of the run and dividing the difference by the difference in the `earth_c` pool from the start of the run to the end, i.e. how much carbon has left the `earth_c` pool.

``` {r hector.define_functions}

tracking_results <- function(inifile, start, stop, tracking, scenarioName) {
  # Establish core, turn on tracking, run core
  core <- newcore(inifile)
  setvar(core, NA, TRACKING_DATE(), tracking, "(unitless)")
  
  reset(core)
  run(core)
  
  # Get results, filter by correct years, add columns containing the amount of carbon by source and the given scenario name
  results <- get_tracking_data(core)
  results <- subset(results, year >= start)
  results <- subset(results, year <= stop)
  results$source_amount <- results$source_fraction * results$pool_value
  results$scenario_name <- scenarioName
  
  return(results)
}

AF_calc <- function(start, results, stop) {
  
  # Isolate the atmos_c pool and earth_c source in start/end years  
  atm_c <- subset(results, results$pool_name == "atmos_c")
  atm_c <- subset(atm_c, atm_c$source_name == "earth_c")
  
  atm_c1 <- subset(atm_c, year == start)
  atm_c2 <- subset(atm_c, year == stop)
  atm_c <- rbind(atm_c1, atm_c2)
  
  # Calculate atmospheric uptake of Earth carbon
  atm_c <- tail(atm_c$source_amount, n = 1) - head(atm_c$source_amount, n = 1)
  # Isolate earth_c pool and source in start/end years
  earth_c_loss <- subset(results, results$pool_name == "earth_c")
  earth_c_loss <- subset(earth_c_loss, earth_c_loss$source_name == "earth_c")
  e_c1 <- subset(earth_c_loss, year == start)
  e_c2 <- subset(earth_c_loss, year == stop)
  earth_c_loss <- rbind(e_c1, e_c2)
  
  # Calculate change in Earth carbon
  earth_c_loss <- head(earth_c_loss$source_amount, n = 1) - tail(earth_c_loss$source_amount, n = 1)
  
  # Airborne fraction calculation
  AF <- atm_c / earth_c_loss
  
  return(AF)
}
```

Next, we will run our functions for the RCP 4.5 scenario between 1750 and 2050 with tracking beginning at the start of the run. We can then find the value of the airborne fraction in any year of the model run. Let's look at how the airborne fraction changes over time from 1950 to 2020 in ten year increments as well as the value in just the year 2020.

``` {r hector.calculations}
rcp45 <- system.file("input", "hector_rcp45.ini", package = "hector")
results <- tracking_results(rcp45, 1750, 2050, 1750, "RCP 4.5")

years <- seq(1950, 2020, by = 10)
airborne_frac <- cbind(make.row.names = years, lapply(years, AF_calc, results, 1750))
colnames(airborne_frac) <- c("year", "AF")

tail(airborne_frac, n = 1)

```

We see that the airborne fraction in 2020 is about 41%, so 41% of anthropogenic emissions remain in the atmosphere. 

We can compare these values to a study published by W. Knorr in 2009.^[Knorr, W. "Is the airborne fraction of anthropogenic CO$_2$ emissions increasing?" Geophysical Research Letters 36 (21). November 7, 2009. http://dx.doi.org/10.1029/2009GL040613] Knorr computes seven values of airborne fractions in 2000 which we can average. We can compute the difference between the maximum value and the minimum to give an error range.

``` {r hector.plot_AF, fig.width = 6, fig.height = 4}
# Knorr avg - values from Knorr 2009
knorr_values <- c(0.45, 0.453, 0.518, 0.468, 0.468, 0.514, 0.449)
knorr_avg <- mean(knorr_values)
knorr_minmax <- max(knorr_values) - min(knorr_values)


names <- c("Knorr avg", years)
values <- c(knorr_avg, as.numeric(airborne_frac[,2]))
error <- c(knorr_minmax, rep(0, times = 8))

df <- data.frame(names, values, error)

ggplot(df) +
  aes(x = names, y = values, fill = names) + 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = values - error, ymax = values + error), width = 0.2)+
  scale_fill_manual(values = c("paleturquoise3", "paleturquoise2", "paleturquoise1", "darkslategray1", "darkslategray2", "darkslategray3", "darkslategray4", "darkslategray", "grey50")) +
  theme(legend.position = "none") +
  labs(x = "Year",
       y = "Airborne fraction over time",
       title = "Airborne Fraction Composition")

```

The airborne fraction found in Hector is towards the lower end of the range of values found by Knorr 2009, which varied from about 45% - 52%. However, we see that the airborne fraction value from Hector remains relatively constant over time at around 40%. This aligns with Knorr's conclusion that the airborne fraction shows no clear trend of increasing or decreasing. Additionally, Hector's value falls within the range of CMIP models, as seen in Jones et al. 2013.^[Jones, C., Robertson, E., Arora, V., Friedlingstein, P., Shevliakova, E., Bopp, L., Brovkin, V., Hajima, T., Kato, E., Kawamiya, M., Liddicoat, S., Lindsay, K., Reick, C. H., Roelandt, C., Segschneider, J., and Tjiputra, J. "Twenty-First-Century Compatible CO2 Emissions and Airborne Fraction Simulated by CMIP5 Earth System Models under Four Representative Concentration Pathways." Journal of Climate, 26 (13) p4398-4413. September 2021. https://doi.org/10.1175/JCLI-D-12-00554.1]

## Conclusion

The carbon tracking feature allows users to identify and trace the composition 
by source of each pool in Hector. This is useful to, for example, calculate the model's airborne fraction. 

Remember to shut down the core.

```{r shutdown}
shutdown(core)
```
