---
title: "Example: using Hector to solve for an emissions pathway"
author: "Robert Link"
date: "2018-10-17"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: solving for an emissions pathway}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This tutorial shows how to use the Hector climate model in R to answer a simple science 
question.  In this case, we'll be looking at solving for the methane emissions required 
over time to produce a constant atmospheric methane concentration from 2000 through 2100.

To run this experiment, you will need to have the `nleqslv` package installed.  This package
will be used to generate successive guesses in the search for an emissions pathway that produces
the target concentration.

## Setup

The `nleqslv` package requires us to supply a function that takes a vector of inputs (i.e., 
emissions by year) and produces a vector of outputs that are the discrepancies between the 
Hector output and the target concentrations.  In order to create such a function we will need to 
set up an interface to Hector.  Since all of our calculations are going to involve years 2000
and beyond, we will run up through 1999.

```{r hector.init}
library('hector')
hector_inifile <- file.path(system.file('input', package='hector'), 'hector_rcp60.ini')
hcore <- newcore(hector_inifile, suppresslogging=TRUE)
run(hcore, 1999)
```

Our target function will take a vector of emission values as input and return a vector of 
discrepancies between the hector outputs and the target output.  We'll make the target concentration
value 1820 ppbv CH4 (a bit higher than the year 2000 value in the baseline scenario).

```{r targfn.init}
f <- function(emiss)
{
    setvar(hcore, 2000:2100, EMISSIONS_CH4(), emiss, "Tg CH4")
    reset(hcore, 1999)
    run(hcore, 2100)
    hout <- fetchvars(hcore, 2000:2100, ATMOSPHERIC_CH4())
    ## return the difference between the target of 1850 (constant) and the actual
    hout$value - 1820
}
```

## Solve for the target emissions

To solve for the target emissions we just pass the target function to `nleqslv`.  We also need an initial guess
for the emissions.  It doesn't really matter what the guess is, so long as it isn't something so large that it 
will cause the model to fail outright.  We will use a constant 300 Tg CH4 as the initial guess.

```{r solve}
x <- rep(300.0, times=101)  # 2000:2001 includes both endpoints
slv <- nleqslv::nleqslv(x, f, method="Broyden")
```

We should check to see that the algorithm really converged.  The discrepancy values from the last iteration
are stored in the `fvec` entry.
```{r chkf}
max(abs(slv$fvec))
```

## Results

The convergence looks ok, so let's plot the results:

```{r plt}
plot(2000:2100, slv$x, ylab="Emissions (Tg CH4)", xlab="year") 
```

The concentration target of 1820 ppbv CH4 that we specified in 2000 is a bit of a jump from what was in the 
scenario we based our run on (the original was about 1811 ppbv), so we have a small spike in emissions in 2000.  After
that the emissions settle down into something smoother.

Once we're done with our Hector instance, we need to make sure to shut it down, so that the memory it is using 
will be freed.
```{r shutdown}
shutdown(hcore)
```

## Conclusion

This simple example shows the kind of thing you can do by combining the Hector R interface with other packages
available in R.  It would be easy enough to add emissions for other gasses or work in terms of forcing or temperature.
Likewise, instead of solving for an equilibrium you could minimize or maximize some function of the emissions (using the
`optim` function), or run one of the many Monte Carlo packages that are available in R.
