---
title: "Land Ocean Warming"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{land_ocean_warming}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE, warning = FALSE, message = FALSE,
    comment = "#>", fig.width = 6, fig.height = 5
)
```

## Introduction

Hector uses DOECLIM^[Kriegler, E. (2005). Imprecise probability analysis for integrated assessment of climate change. Ph.D. thesis. Potsdam URL: Universit t
Potsdam. https://www.pik‐potsdam.de/members/edenh/theses/PhDKriegler.pdf], a diffusion ocean-energy balance, to calculate mean global air temperature from sea-surface and land-surface temperature anomalies. DOECLIM can simulate the different rates of warming over land and ocean surfaces, meaning that the land-ocean warming ratio is an emergent property of Hector. However, users do have the option to provide a land-ocean warming ratio to be used in place of DOECLIM’s internal calculations. This capability enables Hector to emulate the land-ocean warming dynamics of specific Earth System Models. 

The purpose of this vignette is to demonstrate how providing a user-defined land-ocean warming ratio can affect Hector output. 
## Setup

```{r setup}
library(ggplot2)
library(hector)
library(magrittr)

theme_set(theme_bw(base_size = 14))
```


## Run default Hector 

Complete a default run of Hector, one that uses DOECLIM to calculate land and ocean air temperature. 

```{r}
years_to_keep <- 1850:2100
vars_to_keep <- c(GLOBAL_TEMP(), LAND_AIR_TEMP(), OCEAN_AIR_TEMP(), ATMOSPHERIC_CO2())

inifile <- file.path(system.file('input', package = 'hector'), 'hector_ssp245.ini')

core    <- newcore(inifile)
run(core, max(years_to_keep))
out1 <- fetchvars(core, years_to_keep, vars_to_keep)
out1$scenario <- "default"
out1$variable <- paste0(gsub(out1$variable, pattern = "_", replacement = " "), " (", out1$units, ")")
```

```{r}
ggplot(data = out1) + 
    geom_line(aes(year, value)) + 
    facet_wrap("variable", scales = "free") + 
    labs(title = "Default Hector", 
         subtitle = "ssp245", 
         x = NULL) +
    theme(axis.text.x = element_text(angle = 90))
```


## Run Hector with different land ocean warming ratios 

Start by writing a function to help us set the land ocean warming ratio, run hector, and fetch some results.  

```{r}
# Helper function that runs a Hector core with a new land-ocean warming ratio. 
# 
# Args 
#   hc: an active Hector core 
#   value: the land ocean warming ratio to use 
# Return: data frame of Hector results
run_with_lo <- function(hc, value){
    
    # Set the land ocean warming ratio. 
    setvar(hc, NA, LO_WARMING_RATIO(), value, getunits(LO_WARMING_RATIO()))
    
    # Reset and run Hector
    reset(hc)
    run(hc)
    
    # Fetch and format the output. 
    out <- fetchvars(hc, years_to_keep, vars_to_keep)
    out$scenario <- value
    out$variable <- paste0(gsub(out$variable, pattern = "_", replacement = " "), " (", out$units, ")")
    
    return(out)
}


# Make a vector of the land ocean warming ratios to test out. 
lo_to_use <- seq(from = 1, to = 2.5, length.out = 20)

# Apply our helper function to the various land ocean warming ratio values & concatenate the results into a single data frame.
lapply(lo_to_use, run_with_lo, hc = core) %>%
    do.call(what = "rbind") -> 
    hector_lo_results

```


Plot the results 

```{r}
ggplot(data = hector_lo_results) + 
    geom_line(aes(year, value, color = scenario, group = scenario)) + 
    facet_wrap("variable", scales = "free") + 
    labs(title = "User provided land-ocean warming ratio", 
         subtitle = "ssp245", 
         x = NULL, 
         y = NULL) +
    theme(axis.text.x = element_text(angle = 90)) + 
    guides(color=guide_legend(title = "lo ratio"))
```


## Compare default Hector with user provided land ocean warming ratio

```{r}
ggplot() + 
    geom_line(data = hector_lo_results, aes(year, value, color = scenario, group = scenario)) + 
    geom_line(data = out1, aes(year, value), color = "orange", size = 1) + 
    facet_wrap("variable", scales = "free") + 
    labs(title = "User provided land-ocean warming ratio", 
         subtitle = "ssp245", 
         x = NULL, 
         y = NULL) +
    theme(axis.text.x = element_text(angle = 90)) + 
    guides(color=guide_legend(title = "lo ratio"))

```
