---
title: Constraints
subtitle: R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(hector)
library(ggplot2)
theme_set(theme_bw())
```

The Hector model can be run subject to *constraints* that force the model to have a certain behavior. Technically, this means that the model's components output user-provided data as opposed to their own calculations, similar to the [data mode](http://www.cesm.ucar.edu/models/cesm1.0/cesm/cesm_doc_1_0_4/x42.html) of a CESM sub-model. Currently, the available constraints include: 

* Temperature
* Radiative forcing
* GHGs concentrations


Constraints can be applied to Hector within the ini file (command-line & R) or using the `hector` package `setvar` function. 

## Setting Constraints in Hector 

There are two ways to activate a constraint in Hector, through an ini file or with the `setvar` function in R. Here is a generic overview of the two different approaches. See below for a more specific implementation of constraints. 

### Constraints via the ini file

A Hector [input file](InputFiles.html) is broken up into sections such as `[core]`, `[temperature]`, `[carbon-cycle-solver]`, and so on. Within each section, users can define parameter values and pass in input variables including constraints. All of the constraint variables are named with the following pattern `variable_constrain` that points to the csv file where the variables' time series are defined. When the line starts the symbol `;` then the constraint is off, Hector is free running. When the constraint line beings without `;` then the Hector is using the constraint, the user-provided data is used in lieu of Hector's internal calculations. Changes to the constraints in the input files apply to Hector runs executed at the command line or with the R interface. 


### Constraints via `setvar`

With the `hector` R package the `setvar` function can be used to set constraints for a Hector instance. See below for specific examples or run `help("constraints")` or `help("haloconstrain")` for the static R help documentation. 

### Discontinuous constraints

Hector _does not perform any interpolation of constraints_ -- they are applied only in the exact years that they are provided by the user, and all other years are treated (including those immediately before or after constrained years) are treated as unconstrained (typically, emissions-driven).
This means that constraints can abruptly force Hector's climate system into a state that is inconsistent with the previous time step.
Concentrations for the first year of a constraint can be discontinuous compared to the previous year without a constraint.
Concentrations after the last year of a constraint will not necessarily have a discontinuity, but the behavior of the corresponding component and any related components may change after this point.
It is the user's responsibility to make sure that constraints are continuous.


## Current Constraint Capabilities 

### Temperature

Uncommenting (removing the `;`) from the following line in the `[temperature]` section of the [input file](InputFiles.html) will turn the global mean constraint on. 

```
tgav_constrain=csv:path/to/constraint/file.csv
```

Then the model's global mean temperature `Tgav` will follow the contents of this file (which must contain columns titled `Date` and `Tgav_constrain`; a sample file is included in the repository (see `inst/input/tables/tgav_historical.csv`). The model will extrapolate within this data as necessary, but not beyond its endpoints.

Once Hector's internal date passes the last date in the constraint file, the model's temperature component becomes *unconstrained*, except that we do not permit it to suddenly jump to a new temperature. Instead, changes in global radiative forcing (from which temperature change is computed) are applied to the end-of-constraint temperature. For more details, see the detailed comments in the `TemperatureComponent::run` (temperature_component.cpp).

Here is an example of using the R package to implement the global average mean temperature (tgav) constraint. 

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run 
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # the invisible call prevents hector core info from being printed out 

# Extract the output. 
dates_to_keep <- 1850:2100
vars_to_keep <- c(GLOBAL_TEMP())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new temperature values to use as the constraint. 
new_tgav <- free_out[free_out$variable == GLOBAL_TEMP(), ]
new_tgav$value <- new_tgav$value + new_tgav$value * 0.25 # increase global temp by 25%

# Use setvar to set up the the constraint, run hector & fetch the results.  
setvar(core, dates = new_tgav$year, var = TGAV_CONSTRAIN(), values = new_tgav$value, unit = getunits(TGAV_CONSTRAIN()))
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```


Plot the total radiative forcing and global mean temperature. Recall the constraint is only in effect between 1950 and 2050. 

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() + 
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) + 
    geom_line(data = constrained_out, aes(year, value, color = "Constrained RF total"), size = 1,  linetype = 5) + 
    facet_wrap("variable", scales = "free", nrow = 2) + 
    theme(legend.title = element_blank()) + 
    labs(y = NULL, x = "Year")
```


### Radiative forcing

The model's global radiative forcing `Ftot` which drives global temperature changes can be constrained. Free running Hector calculates `Ftot` based on the radiative forcing contributions from GHG concentrations, aerosol emissions, and forcing inputs [UNSURE WORDING] (e.g. land albedo and solar forcing). Using the radiative forcing constraint `Ftot_constrain`, means any parameters and or variables that effect radiative forcing (such as carbon cycle feedback parameters or emissions) will not affect `Ftot` while the constraint is in effect. The model will extrapolate within the `Ftot_constrain` as necessary, but not beyond its endpoints. Once Hector's internal date passes the last date for `Ftot_constrain`, the model's forcing component becomes *unconstrained*.


Uncommenting (removing the `;`) from the following line in the `[forcing]` section of the [input file](InputFiles.html) will turn the radiative forcing constraint on. 

```
;Ftot_constrain=csv:path/to/constraint/file.csv
```

Here is an example of using the R package to implement the radiative forcing constraint.  

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run 
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # the invisible call prevents hector core info from being printed out 

# Extract the output. 
dates_to_keep <- 1850:2100
vars_to_keep <- c(RF_TOTAL(), GLOBAL_TEMP())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new RF values to use as the RF constraint. We are only 
# going to change the RF for a subset of the years. 
new_rf <- free_out[free_out$variable == RF_TOTAL() & free_out$year %in% 1950:2050, ]
new_rf$value <- new_rf$value + 3 # increase the total RF

# Use setvar to set up the the constraint, run hector & fetch the results.  
setvar(core, dates = new_rf$year, var = FTOT_CONSTRAIN(), values = new_rf$value, unit = "W/m2")
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```

Plot the total radiative forcing and global mean temperature. Recall the constraint is only in effect between 1950 and 2050. 

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() + 
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) + 
    geom_line(data = constrained_out, aes(year, value, color = "Constrained RF total"), size = 1,  linetype = 5) + 
    facet_wrap("variable", scales = "free", nrow = 2) + 
    theme(legend.title = element_blank()) + 
    labs(y = NULL, x = "Year")
```


### Atmospheric CO~2~

CO$_2$ concentrations ([CO$_2$], given in ppmv CO$_2$) have the largest effect on total radiative forcing and is affected by carbon-cycle climate interactions. Running Hector with the CO$_2$ concentration constraint turned on not only mimics the default behavior of the Earth System Models which are GHG concentration driven. This capability is also helpful with debugging. With the atmospheric CO$_2$ constraint turned on carbon-cycle and climate feedback parameters ($Q_{10}$ and $\beta$) are essentially ignored.  


Uncommenting (removing the `;`) from the following line in the `[simpleNbox]`section of the [input file](InputFiles.html) will turn the atmospheric CO$_2$ concentration constraint on. 

```
;CO2_constrain=csv:path/to/constraint/file.csv
```


then the model's atmospheric CO$_2$ concentration ([CO$_2$], given in ppmv CO$_2$) will follow the contents of this file (which must include at least two columns titled `Date` and `CO2_constrain`).
Alternatively, Hector's atmospheric CO$_2$ concentration constraint can be set using the `setvars` function and `CO2_CONSTRAIN()`.

Here is an example in R

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run, this is the free running Hector that to be 
# use as a comparison to the constrained Hector. 
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # the invisible call prevents hector core info from being printed out 

# Extract the output. 
dates_to_keep <- 1850:2100
vars_to_keep <- c(ATMOSPHERIC_CO2(), RF_TOTAL(), GLOBAL_TEMP())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new CO2 values to use as the constraint. 
constraint_yrs <- 1950:2050
new_co2 <- seq(from  = 200, to = 400, length.out = length(constraint_yrs))

# Use setvar to set up the the constraint, run hector & fetch the results.  
setvar(core, dates = constraint_yrs, var = CO2_CONSTRAIN(), values = new_co2, unit = getunits(CO2_CONSTRAIN()))
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```

Plot results from the free running Hector with the constrained Hector. 

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() + 
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) + 
    geom_line(data = constrained_out, aes(year, value, color = "x"),
              size = 1,  linetype = 5) + 
    facet_wrap("variable", scales = "free", nrow = 2) + 
    theme(legend.title = element_blank()) + 
    scale_color_discrete(labels=c("Free running", expression(CO[2]~Constrained))) + 
    labs(y = NULL, x = "Year")
```

**How the constraint is implemented in Hector (C++)**

The CO$_2$ constraint is applied at the _end_ of the current time step.
The full sequence of events is as follows:
First, the model solves the current time-step's carbon cycle conditioned on the previous time step's carbon pools, ignoring the CO$_2$ constraint.
Then, if a CO$_2$ constraint is present, Hector calculates the difference between its calculated atmospheric CO$_2$ and the prescribed CO2 constraint.
If the target atmospheric CO$_2$ concentration is lower than the calculated CO$_2$, the excess carbon is transferred from the atmosphere to the deep ocean;
conversely, if the target atmospheric CO$_2$ concentration is greater than the calculated value, the additional carbon is transferred from the deep ocean into the atmosphere.
Finally, Hector records the current atmospheric CO$_2$ concentration (which is now equal to the constraint value); this is the value that is used in the next time step's evaluation of carbon-climate feedbacks (e.g. CO$_2$ fertilization of net primary productivity, surface ocean carbonate chemistry).
In other words, an atmospheric CO$_2$ constraint at time $t$ does not affect carbon-climate feedbacks until $t+1$.

The CO$_2$ constraint need not span the entirety of the Hector simulation, or even be continuous.
At any given time step, Hector will check whether or not a CO$_2$ constraint exists at that time step and only apply the constraint if it is present.
This means that "hybrid" runs are possible, where only some specific time ranges have CO$_2$ constraints while others calculate the atmospheric carbon pool and CO$_2$ concentrations according to Hector's standard [carbon cycle](SimpleNBox.html).

### Non CO~2~ concentrations

Hector can constrain non CO~2~ GHGs (N~2~O, CH~4, and halocarbons) concentrations. These all work in more-or-less the same way as the atmospheric CO$_2$ constraint (above), except that there is no equivalent to the carbon pool adjustment for these gases. 

In the [input file](InputFiles), they are triggered by a line like the following in their corresponding component section:

```
X_constrain=csv:path/to/file.csv
```

As with other CSV inputs, the `/path/to/file.csv` above must include columns for `Date` and `X_constrain`. 

Alternatively, constraints can be set from the R interface using `setvars`. Because implementing these constraints follows the same pattern as the temperature, radiative forcing, and atmospheric CO~2~. Set up an active Hector core, use `setvars` to set the constraint, run, and fetch the output from the Hector core. 


**How the constraint is implemented in Hector (C++) using CH~4~ as an example**

At any given time step, Hector will check if the constraint has a value for that time step (e.g. `CH4_constrain.exists(t)`).
If the constraint exists, Hector will skip all atmospheric chemistry calculates and set the corresponding atmospheric concentration value to the constraint value (e.g. `CH4.set(t, CH4_constrain.get(t)`).
If the constraint does not exist, Hector will proceed with its standard atmospheric chemistry routines for calculating the atmospheric concentration based on emissions and concentrations of other relevant species.
This means that _concentration constraints always take precedence over emissions_; in other words, if a particular time step has both emissions and a concentration constraint, the emissions are ignored.

Note that CH~4~ and N~2~O have special variables for their pre-industrial concentration that could conflict with the constraint. In Hector, _concentration constraints take precedence over pre-industrial concentrations_;
in other words, Hector assumes that if you provide a concentration constraint for the pre-industrial timestep, that is the concentration you want, regardless of the prescribed pre-industrial value. The equations for CH~4~ and N~2~O concentrations to raditive forcing were parameterized with specific pre-industrial concentrations. Users must be cautions of changes to pre-industrial values and how it may affect the radiative forcinig calculations. 
