---
title: Constraints
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(hector)
library(ggplot2)
theme_set(theme_bw())
```

The Hector model can be run subject to *constraints* that force the model to
have a certain behavior. Technically, this means that the model's components output
user-provided data as opposed to their own calculations, similar to the [data mode](http://www.cesm.ucar.edu/models/cesm1.0/cesm/cesm_doc_1_0_4/x42.html) of a
CESM sub-model. Currently, the available constraints include:

* GHG concentrations
* Radiative forcing
* Temperature
* Net Biome Production (NBP)

Constraints can be applied to Hector within the ini file (command-line & R) or
using the Hector package `setvar()` function.

## Setting Constraints in Hector

There are two ways to activate a constraint in Hector: through an ini file or
with the `setvar()` function in R. Here is a generic overview of the two different
approaches. See below for a more specific implementation of constraints.

### Constraints via the ini file

A Hector [input file](InputFiles.html) is broken up into sections such as `[core]`,
`[temperature]`, `[carbon-cycle-solver]`, and so on. Within each section, users
can define parameter values and pass in input variables -- including constraints.
All of the constraint variables are named with the pattern `variable_constrain`
that points to the csv file where the variables' time series are defined. When
the line starts the symbol `;`, the constraint is off and Hector is free running.
When the constraint line beings without `;`, then Hector is using the constraint
and the user-provided data is used in lieu of Hector's internal calculations.
Changes to the constraints in the input files apply to Hector runs executed at
the command line or with the R interface.

### Constraints via `setvar()`

With the Hector R package, the `setvar()` function can be used to set constraints
for a Hector instance. See below for specific examples or run `help("constraints")`
or `help("haloconstrain")` for the static R help documentation.

### Discontinuous constraints

Hector _does not perform any interpolation of constraints_ -- constraints are applied
only in the exact years that they are provided by the user, and all other years
(including those immediately before or after constrained years) are treated as
unconstrained (typically, emissions-driven).
This means that constraints can abruptly force Hector's climate system into a
state that is inconsistent with the previous time step.
Concentrations for the first year of a constraint can be discontinuous compared
to the previous year without a constraint.
Concentrations after the last year of a constraint will not necessarily have a
discontinuity, but the behavior of the corresponding component and any related
components may change after this point.
It is the user's responsibility to make sure that constraints are continuous.


## Current Constraint Capabilities

### Atmospheric CO~2~

CO$_2$ concentrations ([CO$_2$], given in ppmv CO$_2$) have the largest effect on
total radiative forcing and are affected by carbon-cycle climate interactions.
Running Hector with the CO$_2$ concentration constraint turned on mimics the
default behavior of the Earth System Models which are GHG concentration driven.
This capability is also helpful with debugging. With the atmospheric CO$_2$
constraint turned on, carbon-cycle and climate feedback parameters
($Q_{10}$ and $\beta$) are essentially ignored.


Uncommenting (removing the `;`) from the following line in the `[simpleNbox]`
section of the [input file](InputFiles.html) will turn the atmospheric CO$_2$
concentration constraint on.

```
;CO2_constrain=csv:path/to/constraint/file.csv
```

Then the model's atmospheric CO$_2$ concentration ([CO$_2$], given in ppmv CO$_2$)
will follow the contents of this file (which must include at least two columns
titled `Date` and `CO2_constrain`).
Alternatively, Hector's atmospheric CO$_2$ concentration constraint can be set
using the `setvar()` function and `CO2_CONSTRAIN()`.

Here is an example in R.

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run, this is the free running Hector to be
# used as a comparison to the constrained Hector
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # `invisible` prevents Hector core info from being printed out

# Extract the output
dates_to_keep <- 1850:2100
vars_to_keep <- c(ATMOSPHERIC_CO2(), RF_TOTAL(), GLOBAL_TAS())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new CO2 values to use as the constraint
constraint_yrs <- 1950:2050
new_co2 <- seq(from  = 200, to = 400, length.out = length(constraint_yrs))

# Use setvar() to set up the the constraint, reset and run Hector, and fetch the results
setvar(core, dates = constraint_yrs, var = CO2_CONSTRAIN(), values = new_co2,
       unit = getunits(CO2_CONSTRAIN()))
reset(core)
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```

Plot results from the free running Hector with the constrained Hector.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() +
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) +
    geom_line(data = constrained_out, aes(year, value, color = "x"),
              size = 1,  linetype = 5) +
    facet_wrap("variable", scales = "free", nrow = 2) +
    theme(legend.title = element_blank()) +
    scale_color_discrete(labels = c("Free running", expression(CO[2]~Constrained))) +
    labs(y = NULL, x = "Year")
```

**How the constraint is implemented in Hector (C++)**

The CO$_2$ constraint is applied at the _end_ of the current time step.
The full sequence of events is as follows:
First, the model solves the current time-step's carbon cycle conditioned on the
previous time step's carbon pools, ignoring the CO$_2$ constraint.
Then, if a CO$_2$ constraint is present, Hector calculates the difference between
its calculated atmospheric CO$_2$ and the prescribed CO$_2$ constraint.
If the target atmospheric CO$_2$ concentration is lower than the calculated CO$_2$,
the excess carbon is transferred from the atmosphere to the deep ocean;
conversely, if the target atmospheric CO$_2$ concentration is greater than the
calculated value, the additional carbon is transferred from the deep ocean into
the atmosphere.
Finally, Hector records the current atmospheric CO$_2$ concentration (which is now
equal to the constraint value); this is the value that is used in the next time
step's evaluation of carbon-climate feedbacks (e.g. CO$_2$ fertilization of net
primary productivity, surface ocean carbonate chemistry...).
In other words, an atmospheric CO$_2$ constraint at time $t$ does not affect
carbon-climate feedbacks until $t+1$.

The CO$_2$ constraint need not span the entirety of the Hector simulation, or even
be continuous.
At any given time step, Hector will check whether or not a CO$_2$ constraint exists
at that time step and only apply the constraint if it is present.
This means that "hybrid" runs are possible, where only some specific time ranges
have CO$_2$ constraints while others calculate the atmospheric carbon pool and CO$_2$
concentrations according to Hector's standard [carbon cycle](SimpleNBox.html).

### Non CO~2~ concentrations

Hector can constrain non CO$_2$ GHG concentrations (N$_2$O, CH$_4$, and halocarbons).
These all work in more-or-less the same way as the atmospheric CO$_2$ constraint
(above), except that there is no equivalent to the carbon pool adjustment for these gases.

In the [input file](InputFiles), they are triggered by a line like the following
in their corresponding component section:

```
X_constrain=csv:path/to/file.csv
```

As with other CSV inputs, the `/path/to/file.csv` above must include columns for
`Date` and `X_constrain`.

Alternatively, constraints can be set from the R interface using `setvar()`.
Implementing these constraints follows the same pattern as the temperature,
radiative forcing, and atmospheric CO$_2$: Set up an active Hector core, use
`setvar()` to set the constraint, run, and fetch the output from the Hector core.


**How the constraint is implemented in Hector (C++) using CH~4~ as an example**

At any given time step, Hector will check if the constraint has a value for that
time step (e.g. `CH4_constrain.exists(t)`).
If the constraint exists, Hector will skip all atmospheric chemistry calculations
and set the corresponding atmospheric concentration value to the constraint value
(e.g. `CH4.set(t, CH4_constrain.get(t)`).
If the constraint does not exist, Hector will proceed with its standard atmospheric
chemistry routines for calculating the atmospheric concentration based on emissions
and concentrations of other relevant species.
This means that _concentration constraints always take precedence over emissions_;
in other words, if a particular time step has both an emissions and a concentration
constraint, the emissions constraint is ignored.

Note that CH$_4$ and N$_2$O have special variables for their pre-industrial
concentration that could conflict with the constraint. In Hector, _concentration
constraints take precedence over pre-industrial concentrations_;
in other words, Hector assumes that if you provide a concentration constraint for
the pre-industrial timestep, that is the concentration you want, regardless of the
prescribed pre-industrial value. The equations for CH$_4$ and N$_2$O concentrations
to radiative forcing were parameterized with specific pre-industrial concentrations.
Users must be cautions of changes to pre-industrial values and how it may affect
the radiative forcing calculations.

### Radiative forcing

The model's global radiative forcing `Ftot` which drives global temperature changes
can be constrained. Free running Hector calculates `Ftot` based on the radiative
forcing contributions from GHG concentrations, aerosol emissions, and forcing inputs
[UNSURE WORDING] (e.g. land albedo and solar forcing). Using the radiative forcing
constraint `Ftot_constrain` means any parameters and/or variables that affect
radiative forcing (such as carbon cycle feedback parameters or emissions) will not
affect `Ftot` while the constraint is in effect. The model will extrapolate within
the `Ftot_constrain` as necessary, but not beyond its endpoints. Once Hector's
internal date passes the last date for `Ftot_constrain`, the model's forcing
component becomes *unconstrained*.


Uncommenting (removing the `;`) from the following line in the `[forcing]` section
of the [input file](InputFiles.html) will turn the radiative forcing constraint on.

```
;Ftot_constrain=csv:path/to/constraint/file.csv
```

Here is an example of using the R package to implement the radiative forcing constraint.

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # the invisible call prevents Hector core info from being printed out

# Extract the output.
dates_to_keep <- 1850:2100
vars_to_keep <- c(RF_TOTAL(), GLOBAL_TAS())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new RF values to use as the RF constraint. We are only
# going to change the RF for a subset of the years.
new_rf <- free_out[free_out$variable == RF_TOTAL() & free_out$year %in% 1950:2050, ]
new_rf$value <- new_rf$value + 3 # increase the total RF

# Use setvar() to set up the the constraint, reset and run Hector, and fetch the results.
setvar(core, dates = new_rf$year, var = FTOT_CONSTRAIN(), values = new_rf$value, unit = "W/m2")
reset(core)
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```

Plot the total radiative forcing and global mean temperature. Recall the constraint
is only in effect between 1950 and 2050.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() +
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) +
    geom_line(data = constrained_out, aes(year, value, color = "Constrained RF total"),
              size = 1,  linetype = 5) +
    facet_wrap("variable", scales = "free", nrow = 2) +
    theme(legend.title = element_blank()) +
    labs(y = NULL, x = "Year")
```


### Temperature

Uncommenting (removing the `;`) from the following line in the `[temperature]`
section of the [input file](InputFiles.html) will turn the global mean constraint on.

```
tas_constrain=csv:path/to/constraint/file.csv
```

Then the model's global mean temperature `global_tas` will follow the contents of
this file (which must contain columns titled `Date` and `tas_constrain`; a sample
file is included in the repository (see `inst/input/tables/tgav_historical.csv`).
The model will extrapolate within this data as necessary, but not beyond its endpoints.

Once Hector's internal date passes the last date in the constraint file, the model's
temperature component becomes *unconstrained*, except that we do not permit it to
suddenly jump to a new temperature. Instead, changes in global radiative forcing
(from which temperature change is computed) are applied to the end-of-constraint
temperature. For more details, see the detailed comments in the
`TemperatureComponent::run` (temperature_component.cpp).

Here is an example of using the R package to implement the global average mean
temperature (`global_tas`) constraint.

```{r, message=FALSE, warning=FALSE}
# Set up an instance of Hector & run
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # `invisible` prevents Hector core info from being printed

# Extract the output
dates_to_keep <- 1850:2100
vars_to_keep <- c(GLOBAL_TAS())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Make a time series of new temperature values to use as the constraint
new_tgav <- free_out[free_out$variable == GLOBAL_TAS(), ]
new_tgav$value <- new_tgav$value + new_tgav$value * 0.25 # increase global temp by 25%

# Use setvar() to set up the the constraint, reset and run Hector, and fetch the results
setvar(core, dates = new_tgav$year, var = TAS_CONSTRAIN(), values = new_tgav$value,
       unit = getunits(TAS_CONSTRAIN()))
reset(core)
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```


Plot the total radiative forcing and global mean temperature. Recall the constraint
is only in effect between 1950 and 2050.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() +
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) +
    geom_line(data = constrained_out, aes(year, value, color = "Constrained RF total"),
              size = 1,  linetype = 5) +
    facet_wrap("variable", scales = "free", nrow = 2) +
    theme(legend.title = element_blank()) +
    labs(y = NULL, x = "Year")
```

### NBP

The NBP constraint can be activated via the ini file by adding the line

```
NBP_constrain=csv:path/to/constraint/file.csv
```

to the `[simpleNbox]` section of the [input file](InputFiles.html), where the
constraint csv file has columns titled `Date` and `NBP_constrain` similar to the
sample file is included in the repository (see `inst/input/tables/tgav_historical.csv`).

Here is an example implemented in the R interface

```{r}
constraint_yrs <- 2001:2010
constraint_values <- c(0.1, 0.2, 0.3, -0.4, -0.5, -0.6, 0.7, 0.8, 0.9, 1.0)

ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245")
invisible(run(core)) # `invisible` prevents Hector core info from being printed out

# Extract the output
dates_to_keep <- 1850:2100
vars_to_keep <- c(ATMOSPHERIC_CO2(), NBP(), NPP(), RH())
free_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)

# Use setvar() to set up the the constraint, reset and run Hector, and fetch the results
setvar(core, dates = constraint_yrs, var = NBP_CONSTRAIN(), values = constraint_values,
       unit = getunits(NBP_CONSTRAIN()))
reset(core)
invisible(run(core))
constrained_out <- fetchvars(core, 1850:2100, vars = vars_to_keep)
```

Plot results from the free running Hector with the constrained Hector.

```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplot() +
    geom_line(data = free_out, aes(year, value, color = "Free running"), size = 1) +
    geom_line(data = constrained_out, aes(year, value, color = "x"),
              size = 1,  linetype = 5) +
    facet_wrap("variable", scales = "free", nrow = 2) +
    theme(legend.title = element_blank()) +
    scale_color_discrete(labels = c("Free running", expression(CO[2]~Constrained))) +
    labs(y = NULL, x = "Year")
```

The model hits the constraint exactly by raising RH and lowering NPP.
Because of this, atmospheric CO$_2$ rises rapidly during the constraint period,
as we've basically shut down the land C sink. As a result, when the model exits
the constraint period, NPP shoots up very high (because the atmosphere is saturated
with CO$_2$).
In a similar vein, the suppressed NPP means, once outside of the constraint, that
RH is low and thus the land becomes a large sink as it draws down the excess
atmospheric CO$_2$.
