---
title: Hector input files
---

# General information

When Hector is invoked from the command line, it takes a single parameter: the name of the primary input file to read. This file, typically stored in the `input/` directory, contains configuration and parameter data controlling the model's run, output, and behavior.

The primary input file is a text file with *sections*, *properties*, and *values*. For example, the beginning of the [SSP245 file](/tree/master/inst/input/hector_ssp245.ini) included in the repository looks like this:

```
; Config file for hector model: SSP245
;------------------------------------------------------------------------
[core]
run_name=ssp245
```

Here, comments start with a semicolon (`;`); *section names* (`core`) are enclosed in square brackets; and a *property* (`run_name`) is assigned a *value* (`ssp245`). Properties are guaranteed to be unique within a particular section, and all sections match one-to-one with a model [component](ComponentAPI.html).


## Table of model parameters

``` {r table, echo = FALSE, warning = FALSE}
library(knitr)
library(kableExtra)

input <- hector::inputstable
colnames(input) <- c("Section", "Parameter", "Biome specific", "Time variant",
                     "Required", "Default", "Units", "Description")

params <- subset(input, input$`Time variant` == "n")
v_params <- subset(input, input$`Time variant` == "y")
biome_spec <- subset(input, input$`Biome specific` == "y")
colnames(v_params)[2] <- "Input variable"

rownames(params) <- NULL
rownames(v_params) <- NULL
rownames(biome_spec) <-  NULL

kbl(params) %>%
    kable_styling(fixed_thead = TRUE,
                  bootstrap_options = "hover") %>%
      scroll_box(height = "400px")
```

## Table of biome-specific inputs

It is possible to run Hector with multiple distinct [biomes](multiple-biomes.html).
The following parameters are specific to a particular biome.

``` {r biomes, echo = FALSE}
kbl(biome_spec) %>%
    kable_styling(fixed_thead = TRUE,
                  bootstrap_options = "hover") %>%
      scroll_box(height = "400px")

```


## Table of variable inputs (with a time index)

There are two Hector-specific wrinkles to the INI property-value assignment scheme.
First, some properties have a time index that's required when assigning:

```
Ftalbedo[1750]=0.0
Ftalbedo[1950]=-0.2
```

More commonly, such [time series](TimeSeries.html) are assigned by reading data
in from a comma-separated value text file:

```
lucEmissions=csv:input/emissions/ssp245_emissions.csv
```

In this case, the `input/emissions/ssp245_emissions.csv` file is parsed and the
data in column `lucEmissions` are assigned to the model variable, with a required
`Date` column providing the time index. Note that a "(csv)" entry in the `default`
column indicates that the values are passed in via a path to a csv file.

``` {r timetable, echo = FALSE}
kbl(v_params) %>%
    kable_styling(fixed_thead = TRUE,
                  bootstrap_options = "hover") %>%
      scroll_box(height = "400px")
```

[gsl-ode]: https://www.gnu.org/software/gsl/manual/html_node/Ordinary-Differential-Equations.html#Ordinary-Differential-Equations
