---
title: "An Introduction to Permafrost in Hector"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Introduction to Permafrost in Hector}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Hector version 3 introduces a new capability to the model: permafrost thaw.
The is important because enormous amounts of organic carbon are stored in
high-latitude permafrost soils, frozen pools that can destabilized, thawed,
and released to the atmosphere by climatic warming.
This process has the potential to significantly accelerate climate
change [(Schuur et al. 2015)](https://www.nature.com/articles/nature14338).

**It is important to note that (1) permafrost is currently OFF by default,
and (2) the model's permafrost dynamics have not been fully checked against
the original implementation [(Woodard et al. 2021)](https://gmd.copernicus.org/articles/14/4751/2021/). As a result, this
is currently an EXPERIMENTAL feature that is not 
ready for scientific simulations.**

> Woodard, D. L., Shiklomanov, A. N., Kravitz, B., Hartin, C., Bond-Lamberty, B. (2021). A permafrost implementation in the simple carbon-climate model Hector v.2.3pf. Geosci. Model Dev., 14, 4751–4767. https://doi.org/10.5194/gmd-14-4751-2021

### Setup

Start by setting up a default Hector core and running it:

```{r setup}
library(hector)
ini <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini, name = "ssp245-default")
```

As noted above, currently by default the model is initialized with no permafrost carbon:

```{r}
fetchvars(core, dates = NA, vars = PERMAFROST_C())
```

Run the default core and save its outputs:

```{r}
invisible(run(core))

# Output variables we want to retrieve
climate_vars <- c(ATMOSPHERIC_CO2(), GLOBAL_TAS(), RF_CO2(), RF_TOTAL())
permafrost_vars <- c(PERMAFROST_C(), THAWEDP_C(), RH(), CONCENTRATIONS_CH4())

run1 <- fetchvars(core, dates = 1745:2300, vars = c(climate_vars, permafrost_vars))
shutdown(core)
```

### Enabling permafrost

Set up a new core and set its initial permafrost carbon to 865 Pg C,
the value used by Woodard et al. (2021):

```{r}
core <- newcore(ini, name = "ssp245-permafrost")
setvar(core, 0, PERMAFROST_C(), 865, "Pg C")
invisible(run(core))

run2 <- fetchvars(core, dates = 1745:2300, vars = c(climate_vars, permafrost_vars))
```

We can check and confirm that permafrost was lost to the atmosphere as the climate warmed...

```{r}
fetchvars(core, 2300, PERMAFROST_C())
shutdown(core)
```

...leading to earth system and climate differences:

```{r, fig.width=7, fig.height=5}
results <- rbind(run1, run2)

library(ggplot2)
ggplot(subset(results, variable %in% climate_vars), 
       aes(year, value, color = scenario)) + 
    geom_line() +
    facet_wrap(~variable, scales = "free")
```

### A closer look at permafrost dynamics

The thaw rate is controlled by biome-specific air temperature, and thawed
permafrost slowly decomposes into both CH$_4$ and CO$_2$. We can look at some of
these effects in more detail:

```{r, fig.width=7, fig.height=5}
ggplot(subset(results, variable %in% permafrost_vars), 
       aes(year, value, color = scenario)) + 
    geom_line() +
    facet_wrap(~variable, scales = "free")
```

Here we can see how permafrost thaws and then, in the 22nd century, slightly
refreezes; the thawed permafrost pool peaks in the second half of the 20th
century and then declines; and both atmospheric methane and total heterotrophic
respiration are significantly higher
in the permafrost than non-permafrost simulations.

### Controlling the behavior of permafrost thaw

As described by Woodard et al. (2021), the relationship between air temperature
and thaw fraction is controlled by two parameters that are calibrated to 
CMIP6 data:

```{r}
core <- newcore(ini, name = "ssp245-permafrost")
setvar(core, 0, PERMAFROST_C(), 865, "Pg C")
fetchvars(core, NA, c(PERMAFROST_MU(), PERMAFROST_SIGMA()))
```

We can test the effect of varying these parameters:

```{r, fig.width=7, fig.height=5}
run_fn <- function(value, core, variable) {
    setvar(core, NA, variable, value, "degC")
    reset(core)
    run(core)
    out <- fetchvars(core, 2100:2300, vars = GLOBAL_TAS())
    out$varying <- variable
    out$param_value <- value
    out
}

# Run basic parameter sensitivity test
results_mu <- lapply(seq(1.4, 2.1, length.out = 6), run_fn, core, PERMAFROST_MU())
results_sigma <- lapply(seq(0.7, 1.2, length.out = 6), run_fn, core, PERMAFROST_SIGMA())

# Combine results and plot
results <- do.call("rbind", c(results_mu, results_sigma))

default <- subset(run2, variable == GLOBAL_TAS() & year %in% 2100:2300)

ggplot(results, aes(year, value, color = param_value, group = paste(varying, param_value))) +
    geom_line(linetype = 2) +
    ylab(GLOBAL_TAS()) +
    facet_wrap(~varying) +
    # add a thick black line to show default behavior
    annotate(geom = "line", x = default$year, y = default$value, linewidth = 1.5)
```

Hector's permafrost capability is planned to be fully science-ready in v3.2.
